### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Start `ed-notes` doc. Update references, fix typos.

---
title: ED2 model notes
author: Alexey Shiklomanov
---

# ED2 switches

## Finite canopy radius (`CROWN_MOD`)

In `src/dynamics/radiate_driver.f90:L550`:
If disabled, set canopy crown area for all patches to 1.
If enabled, set finite crown areas for each cohort.

Crown area used in some of the following places:

- Canopy structural dynamics (`dynamics/canopy_struct_dynamics.f90:L470`)
- Plant hydrology (`dynamics/plant_hydro.f90:L145,204,240,447,626`)

Crown area comes from @dietze_2008_capturing.
Crown area calculated by allometry in `utils/allometry.f90:L633` (inside subroutine `area_indices`), as `min(1, nplant * dbh2ca(dbh, hite, sla, pft))`.
`dbh2ca` is defined in `utils/allometry.f90:L373` as:

```
ca = biCa[pft] * dbh ^ b2Ca[pft]
```

But, note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):

```
loclai = sla * size2bl(dbh, hite, pft)
ca = min(loclai, ca)
```

where `size2bl` is the leaf biomass allometry (so leaf biomass; therefore, `sla * size2bl` gives LAI).

Crown area set or calculated in:

- `dynamics/strcutral_growth.f90:L669,929`

## Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)

## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)