### 2019-01-31 analysis/paper/ed-notes.Rmd
PAPER: Add notes on finite canopy radius in ED

---
title: ED2 model notes
author: Alexey Shiklomanov
---

# ED2 switches

## Finite canopy radius (`CROWN_MOD`)

### Calculating crown area

Crown area comes from @dietze_2008_capturing.
Crown area calculated by allometry in `utils/allometry.f90:L633` (inside subroutine `area_indices`), as `min(1, nplant * dbh2ca(dbh, hite, sla, pft))`.
`dbh2ca` is defined in `utils/allometry.f90:L373` as:

```
ca = b1Ca[pft] * dbh ^ b2Ca[pft]
```

Note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):

```
loclai = sla * size2bl(dbh, hite, pft)
ca = min(loclai, ca)
```

where `size2bl` is the leaf biomass allometry (so leaf biomass; therefore, `sla * size2bl` gives LAI).

The allometry coefficients `b1Ca` and `b2Ca` have the following defaults (from `init/ed_params.f90:L3055`):

- `b1Ca`
    - 5:13 (C3 grass and temperate trees) -- `2.490154` (constant)
    - All others -- `exp(-1.853) * exp(b1Ht(pft)) ** 1.888`; log-linear relationship with height allometry
- `b2Ca`
    - 5:13 -- `0.8068806` (constant)
    - All others -- `b2Ht(pft) * 1.888`; linear relationship with height allometry

Note that only tropical PFTs are affected by the allometry model (`IALLOM`); coefficient adjustments are wrapped in `if (is_tropical(pft))`.

If two cohorts are fused, their crown areas are summed, up to a maximum of 1 (in `utils/fuse_fiss_utils.f90:L2932`).

### Radiative transfer

In `src/dynamics/radiate_driver.f90:L550`:
If disabled, set canopy crown area for all patches to 1.
If enabled, set finite crown areas for each cohort.

Crown area array (for each cohort) is passed to the various radiative transfer models (see `ICANRAD` parameter).
Inside the RTMs, it is the `cai` variable.

The following sections focus on the contributions of finite canopy radius to radiative transfer calculations.
For the full descriptions, consult the "Canopy radiative transfer" module section.

#### Two-stream

In the two-stream model, crown area is used to correct `mu_bar` (average inverse optical depth per unit leaf and stem area) for crown area (resulting variable `mu`):

```
mu = -etai / log(1 - cai + cai * exp(-etai / (cai * mu_bar)))
```

where `etai` is the total area index (TAI = LAI * clumping factor + WAI) of the layer.

(If `cai = 1`, this simplifies to just `mu_bar`).

In the longwave (thermal), crown area is also used as a linear scaling of the blackbody emissivity.

In the shortwave two-stream model, the crown area correction also happens for direct radiation (`mu0`):

```
mu0 = -etai / log(1 - cai + cai * exp((-proj_area * etai) / (cai * czen)))
```

where `czen` is the cosine of the solar zenith angle (radians) and `proj_area` is the projected area, calculated as a function of leaf orientation angle (`orient`):

```
phi1 = 0.5 - orient * (0.633 + 0.33 * orient)
phi2 = 0.877 * (1 - 2 * phi1)
proj_area = phi1 + phi2 * czen
```

These calculations follow CLM 4.5 [@clm45_note] (section 3.1, eq. 3.3), which in turn is based on @sellers_1985_canopy.
This is described in more detail in the canopy radiative transfer section.

#### Multiple-scatter

In the multiple-scatter model [@zhao_2005_multiple; @zhao_2006_modeling], crown area is used calculate the exposed total area index (TAI) of each layer (`locetai`):

```
locetai = etai / cai
```

Crown area is also used to adjust the layer transmittance coefficient for beam radiation (`tau_beam`):

```
lambda = proj_area / mu
tau_beam = (1 - cai) + cai * exp(-lambda * locetai)
```

A similar correction is applied to diffuse radiation, but the contribution from each hemisphere has to be integrated.
This uses the analytical solution for that integral.

```
d1 = phi1 * locetai
d2 = phi2 * locetai
tau_diff = (1 - cai) - cai * exp(-d1 * d2) * (d1^2 * exp(d1) * eifun(-d1) + (d1 - 1))
```

where `eifun` is the exponential integral function.

### Plant hydraulics

Source: `dynamcs/plant_hydro.f90`.

Subroutine `plant_hydro_driver`:
Crown area for each cohort calculated as `crown_area / nplant`, then passed to `calc_plant_water_flux` (as variable `crown_area`).

Calculate the root area index (RAI) at each depth:

```
RAI = broot_d * SRA * root_frac / (4 * crown_area)
```

where SRA is the PFT specific root area parameter, and `root_frac` is the fraction of roots in that layer (function of PFT parameter `root_beta` and depth).

Calculate the soil-root water conductance at each depth:

```
gw_cond = (4 * crown_area) * soil_cond * sqrt(RAI) / (pi * dslz)
```

where `soil_cond` is soil conductivity at that depth and `dslz` is the soil layer thickness.
This is based on @katul_2003_relationship.

### Canopy turbulence

In file `dynamics/canopy_struct_dynamics.f90`, subroutine `canopy_turbulence`.

Crown area used in calculation of the wind profile extinction coefficients.

```
#-- dynamics/canopy_struct_dynamics.f90:L470 --
extinct_full = crown_area * exp(-0.25 * lai / crown_area) + (1 - crown_area)
extinct_full = crown_area * exp(-0.5 * lai / crown_area) + (1 - crown_area)
veg_wind = max(ugbmin, uh[i - 1] * extinct_half)
uh[i] = uh[i - 1] * extinct_full
```

Note that there is a copy of the same calculation performed in double precision (subroutine `canopy_turbulence8`) later in the file.

Crown area is also used to calculate the open canopy fraction of each cohort by subroutine `update_patch_derived_props` (file `utils/update_derived_props.f90:L43`).
That subroutine is in turn called by subroutine `structural_growth_eq_0` (file `dynamics/structural_growth.f90`).
This open canopy fraction is used to calculate the canopy roughness (`rough`).

```
rough = snow_rough * snowfac_can +
    (soil_rough * opencan_frac + veg_rough * (1 - opencan_frac)) *
    (1 - snowfac_can)
```

Open canopy fraction is also used to calculate the net ground conductance (`ggnet`):

```
ggnet = ggbare * ggveg / (ggveg + ggbare * (1 - opencan_frac))
```

## Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)

## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)