### 2019-03-25 analysis/paper/ed-notes.Rmd
Move old ED module notes from paper to notes

---
title: ED2 model notes
author: Alexey Shiklomanov
---

# ED2 switches

## Finite canopy radius (`CROWN_MOD`)

### Calculating crown area

Crown area comes from @dietze_2008_capturing.
Crown area calculated by allometry in `utils/allometry.f90:L633` (inside subroutine `area_indices`), as `min(1, nplant * dbh2ca(dbh, hite, sla, pft))`.
`dbh2ca` is defined in `utils/allometry.f90:L373` as:

```
ca = b1Ca[pft] * dbh ^ b2Ca[pft]
```

Note that crown area is restricted such that "Local LAI / Crown area" is never less than one (i.e you cannot have more crown area than leaf area index -- your maximum crown area is the leaf area index):

```
loclai = sla * size2bl(dbh, hite, pft)
ca = min(loclai, ca)
```

where `size2bl` is the leaf biomass allometry (so leaf biomass; therefore, `sla * size2bl` gives LAI).

The allometry coefficients `b1Ca` and `b2Ca` have the following defaults (from `init/ed_params.f90:L3055`):

- `b1Ca`
    - 5:13 (C3 grass and temperate trees) -- `2.490154` (constant)
    - All others -- `exp(-1.853) * exp(b1Ht(pft)) ** 1.888`; log-linear relationship with height allometry
- `b2Ca`
    - 5:13 -- `0.8068806` (constant)
    - All others -- `b2Ht(pft) * 1.888`; linear relationship with height allometry

Note that only tropical PFTs are affected by the allometry model (`IALLOM`); coefficient adjustments are wrapped in `if (is_tropical(pft))`.

If two cohorts are fused, their crown areas are summed, up to a maximum of 1 (in `utils/fuse_fiss_utils.f90:L2932`).

### Radiative transfer

In `src/dynamics/radiate_driver.f90:L550`:
If disabled, set canopy crown area for all patches to 1.
If enabled, set finite crown areas for each cohort.

Crown area array (for each cohort) is passed to the various radiative transfer models (see `ICANRAD` parameter).
Inside the RTMs, it is the `cai` variable.

The following sections focus on the contributions of finite canopy radius to radiative transfer calculations.
For the full descriptions, consult the "Canopy radiative transfer" module section.

#### Two-stream

In the two-stream model, crown area is used to correct `mu_bar` (average inverse optical depth per unit leaf and stem area) for crown area (resulting variable `mu`):

```
mu = -etai / log(1 - cai + cai * exp(-etai / (cai * mu_bar)))
```

where `etai` is the total area index (TAI = LAI * clumping factor + WAI) of the layer.

(If `cai = 1`, this simplifies to just `mu_bar`).

In the longwave (thermal), crown area is also used as a linear scaling of the blackbody emissivity.

In the shortwave two-stream model, the crown area correction also happens for direct radiation (`mu0`):

```
mu0 = -etai / log(1 - cai + cai * exp((-proj_area * etai) / (cai * czen)))
```

where `czen` is the cosine of the solar zenith angle (radians) and `proj_area` is the projected area, calculated as a function of leaf orientation angle (`orient`):

```
phi1 = 0.5 - orient * (0.633 + 0.33 * orient)
phi2 = 0.877 * (1 - 2 * phi1)
proj_area = phi1 + phi2 * czen
```

These calculations follow CLM 4.5 [@clm45_note] (section 3.1, eq. 3.3), which in turn is based on @sellers_1985_canopy.
This is described in more detail in the canopy radiative transfer section.

#### Multiple-scatter

In the multiple-scatter model [@zhao_2005_multiple; @zhao_2006_modeling], crown area is used calculate the exposed total area index (TAI) of each layer (`locetai`):

```
locetai = etai / cai
```

Crown area is also used to adjust the layer transmittance coefficient for beam radiation (`tau_beam`):

```
lambda = proj_area / mu
tau_beam = (1 - cai) + cai * exp(-lambda * locetai)
```

A similar correction is applied to diffuse radiation, but the contribution from each hemisphere has to be integrated.
This uses the analytical solution for that integral.

```
d1 = phi1 * locetai
d2 = phi2 * locetai
tau_diff = (1 - cai) - cai * exp(-d1 * d2) * (d1^2 * exp(d1) * eifun(-d1) + (d1 - 1))
```

where `eifun` is the exponential integral function.

### Plant hydraulics

Source: `dynamcs/plant_hydro.f90`.

Subroutine `plant_hydro_driver`:
Crown area for each cohort calculated as `crown_area / nplant`, then passed to `calc_plant_water_flux` (as variable `crown_area`).

Calculate the root area index (RAI) at each depth:

```
RAI = broot_d * SRA * root_frac / (4 * crown_area)
```

where SRA is the PFT specific root area parameter, and `root_frac` is the fraction of roots in that layer (function of PFT parameter `root_beta` and depth).

Calculate the soil-root water conductance at each depth:

```
gw_cond = (4 * crown_area) * soil_cond * sqrt(RAI) / (pi * dslz)
```

where `soil_cond` is soil conductivity at that depth and `dslz` is the soil layer thickness.
This is based on @katul_2003_relationship.

### Canopy turbulence

In file `dynamics/canopy_struct_dynamics.f90`, subroutine `canopy_turbulence`.

Crown area used in calculation of the wind profile extinction coefficients.

```
#-- dynamics/canopy_struct_dynamics.f90:L470 --
extinct_full = crown_area * exp(-0.25 * lai / crown_area) + (1 - crown_area)
extinct_full = crown_area * exp(-0.5 * lai / crown_area) + (1 - crown_area)
veg_wind = max(ugbmin, uh[i - 1] * extinct_half)
uh[i] = uh[i - 1] * extinct_full
```

Note that there is a copy of the same calculation performed in double precision (subroutine `canopy_turbulence8`) later in the file.

Crown area is also used to calculate the open canopy fraction of each cohort by subroutine `update_patch_derived_props` (file `utils/update_derived_props.f90:L43`).
That subroutine is in turn called by subroutine `structural_growth_eq_0` (file `dynamics/structural_growth.f90`).
This open canopy fraction is used to calculate the canopy roughness (`rough`).

```
rough = snow_rough * snowfac_can +
    (soil_rough * opencan_frac + veg_rough * (1 - opencan_frac)) *
    (1 - snowfac_can)
```

Open canopy fraction is also used to calculate the net ground conductance (`ggnet`):

```
ggnet = ggbare * ggveg / (ggveg + ggbare * (1 - opencan_frac))
```

## Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)

### Calculating N supply and demand for live tissue growth

Basically, N limitation places a limit on the fraction of stomata that are open (due to N limitation; `cpatch%fsn`).

Calculating changes in live biomass (subroutines `dbalive_dt` and `dbalive_dt_eq_0` in `dynamics/growth_balive.f90`).

If plant limitation enabled, call subroutine `potential_N_uptake` to see what would happen if stomata were open.
This sets the value of potential N uptake (`N_uptake_pot`).

If no limitation, set fraction of open stomata (from N limitation; `fsn`) to 1.
Otherwise, calculate N limitation factor.

```
#- dynamics/growth_balive.f90:L293
nitrogen_supply = plant_N_supply_scale * broot * mineralized_soil_N
fsn = nitrogen_supply / (nitrogen_supply + N_uptake_pot)
```

### Initialization from near bare ground

Subroutines `near_bare_ground_init` and `near_barea_ground_big_leaf_init` (in `init/ed_nbg_init.f90`):

If there is N limitation, initialize as follows:

- Fast soil C (`fast_soil_C`): 0.2
- Slow soil C (`slow_soil_C`): 0.01
- Structural soil C (`structural_soil_C`): 10.0
- Structural soil lignin (`structural_soil_L`): 10.0
- Mineralized soil N (`mineralized_soil_N`): 1.0
- Fast soil N (`fast_soil_N`): 1.0

If no N limitation, set all of these to zero.

## Nitrogen limitation of decomposition (`N_DECOMP_LIM`)

Calculation of N immobilization factor (subroutine `resp_f_decomp`, in `dynamics/soil_respiration.f90`).

N decomposition limitation affects the decomposition fraction (`csite%f_decomp`).
If no limitation, this is set to 1.

If limitation, calculate similarly to plant (supply relative to demand):

```
#-- dynamics/soil_respiration.f90:L226
Lc = exp(-3 * structural_soil_L / structural_soil_C)

N_immobilization_demand = A_decomp * Lc * decay_rate_stsc * structural_soil_C *
  ((1 - r_stsc) / c2n_slow - 1 / c2n_structural)

f_decomp = N_immobil_supply_scale * mineralized_soil_N / 
  (N_immobilization_demand + N_immobil_supply_scale * mineralized_soil_N)
```

Whether N decomposition limitation is enabled also affects decay rate of slow carbon pool if using `DECOMP_SCHEME = 2`.
With N limitation, decay rate is 100.2 year^-1^; otherwise, 0.2 year^-1^.

# Old notes

## ED modules

NOTE: Not a real section; just brainstorming notes for now
Details about these sub-models will probably makes their way into the methods and/or appendices.

### High priority

These are switches that I understand (more-or-less), and that I think will be most relevant to FoRTe.

- Finite canopy radius (`CROWN_MOD`)
    - Options:
        - (0; default): Complete shading
        - (1): Finite canopy radius (Dietze 2008)
    - See current hypotheses
    - This model is based on crown area allometries from @dietze_2008_capturing
- Nitrogen limitation of photosynthesis (`N_PLANT_LIM`)
    - Options:
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Turning this on _should_ favor mid-successional species, and substantially increase uncertainty.
- Nitrogen limitation of decomposition (`N_DECOMP_LIM`)
    - Options:
        - (0, default): No limitation
        - (1) Limitation based on internal model (TODO: what is it?)
    - Same as above

### Medium priority

These are switches that may be relevant to FoRTE, but may be harder to pull off.

- Big leaf model (no demographics) (`IBIGLEAF`)
    - Simplistic case
    - Only one PFT per patch (by definition). Can ask: Does it get the right one?
    - Compare fluxes against other cases.
- Canopy radiative transfer model (`ICANRAD`)
    - Options:
        - (0, deprecated?) Original two-stream model (Medvigy 2006)
        - (1) Multiple-scattering model [@zhao_2006_modeling; @zhao_2005_multiple]
        - (2, default) Updated two-stream model [@liou_2002]
    - In theory, multiple-scattering model is more detailed and should give "better" results. Open question as to how that affects predictions.
    - Good candidate for exploratory analysis. I already implemented the two-stream model in R -- could be a good idea to implement the multi-scatter model as well and compare their radiation profiles.
- Allometry (`IALLOM`)
    - Options:
        - (0, default): Original ED2
        - (1)
            - Structural biomass parameterized so total AGB similar to @baker_2004_variation, eq. 2. `Balive` similar to ED2.
            - Root depth configured so root depth is 5m for canopy trees and 0.5 m for grasses/seedlings
            - Crown area as in @poorter_2006_architecture, with maximum crown area
        - (2) Like (1), but...
            - Height to DBH allometry from @poorter_2006_architecture
            - `Balive` has extra allometric equations (see @cole_2006_allometric and @calvo-alvarado_2008_allometric)
        - (3) Like (2), but tropical leaf biomass estimated from @lescure_1983_la, and PFT differences in leaf biomass based on SLA rather than wood density. Also, split leaf allometry into small (sapling) and large (tree) classes, as in @cole_2006_allometric
- Temperature variation of physiology (`IPHYSIOL`)
    - Options:
        - (0) Arrhenius function, `Gamma*` based on Foley et al. (1996) (TODO: REF)
        - (1) Arrhenius, `Gamma*` based on Michaelis-Mentel coefficients for CO2 and O2, as in @farquhar_1980_biochemical
        - (2; default) Q10 equations and parameters of @collatz_1991_physiological, with MM coefficients. High/low temperature correction in @moorcroft_2001_method
        - (3) As (2), but `Gamma*` as in @farquhar_1980_biochemical and CLM.
    - Need to do some exploratory analysis here. Grab the relevant equations and coefficients and plot them (outside of ED) to compare behavior.
        - Possible angle: More important at temperature extremes?
    - Consult @rogers_2016_roadmap for inspiration
- Soil decomposition temperature sensitivity (`DECOMP_SCHEME`)
    - Options:
        - (0; default): Exponential
        - (1) Lloyd and Taylor (1994)
    - (1) requires additional parameters, so should increase parameteric uncertainty.
    - Hypothesis: Affects overall GPP/R~E~ balance, but not successional dynamics.
- Water limitation of photosynthesis (`H2O_PLANT_LIM`)
    - Options:
        - (0) No limitation.
        - (1; default): Supply is total soil water above wilting point integrated across all levels in rooting zone.
        - (2): Supply is total soil water at field capacity minus wilting point, scaled by "wilting factor":
          ```
          # k -- Soil layer
          # z -- Layer depth
          # psi -- Matric potential (for layer, or at wp or fc)
          # wp -- wilting point
          # fc -- field capacity
          (psi(k) - (H - z(k)) - psi_wp) / (psi_fc - psi_wp)
          ```
    - Sensitivity should depend on site climate and soil conditions.
    - UMBS are pretty wet sites, so shouldn't expect a huge effect.
- Quantum yield model (`QUANTUM_EFFICIENCY_T`)
    - Options:
        - (0, default): Constant quantum efficiency
        - (1): Quantum efficiency varies with temperature, based on Ehleringer (1978) (TODO: Get ref)
    - Uncertainty should increase with (1) assuming parameters are tunable. 
    - Depends on where UMBS conditions fall within the range of temperatures. Good candidate for exploratory analysis outside of ED.
- Structural growth (`ISTRUCT_GROWTH_SCHEME`)
    - Options:
        - (0; default) Use all `bstorage` (NSC?)
        - (1) Reserve `bstorage` to re-flush canopy and fine roots once before calculating structural growth.
    - (1) is a more conservative strategy, so should reduce the growth and relative dominance of early-successional PFTs. In particular, growth rate should decline more with size (because more LAI and fine roots means more reserves needed).
- Stomatal conductance model (`ISTOMATA_SCHEME`)
    - Options:
        - (0, default): Leuning
        - (1) Katul optimization model (see Xu et al. 2016 New Phyt.)
    - Need to read more about this...
- Trait plasticity (`TRAIT_PLASTICITY_SCHEME`)
    - Options:
        - (0, default): Static traits
        - (1) Vm0 (SLA) decrease (increase) with shading, based on Lloyd et al. (2010 Biogeosciences). Updated yearly.
        - (2) Same as (1), but updated monthly
        - (-1) Same as 1, but adjust SLA using height
        - (-2) Same as (2), but adjust SLA using height
    - If trait plasticity is optimal, this should reduce (?) the relative dominance of early-successional PFTs.
    - Probably introduces considerable predictive uncertainty, given new parameters for plasticity models.
- Carbon balance contribution to mortality (`IDDMORT_SCHEME`)
    - Options:
        - (0, default): Carbon balance in terms of fluxes only
        - (1) Carbon balance is offset by storage pool (i.e. carbon storage is an additional buffer against mortality)
    - Additional protection from (1) should reduce mortality for all trees, so will probably reinforce dominance of early successional trees?
    - But likely to interact with other schemes, especially structural growth, which determines size of NSC reserves
- Percolation and infiltration (`IPERCOL`)
    - Options:
        - (0) Constant soil conductivity. Temporary surface water (only exists if top soil layer is saturated) exceeding 1:9 liquid:ice ratio shed via percolation.
        - (1, default) Constant soil conductivity, with percolation model from Anderson (1976 NOAA technical report NWS 19 TODO). Temporary surface water can exist after heavy rain event even if soil doesn't saturate.
        - (2) Like (1), but soil conductivity decreases with depth even under constant soil moisture
    - Presumably, not a huge impact at UMBS because water shouldn't be an issue.
    - See also `RUNOFF_TIME` (e-folding lifetime of temporary surface water)
- Carbon stress scheme (`CBR_SCHEME`)
    - Options:
        - (0, default): Single stress, with maximum carbon balance determined at full sun and no moisture
        - (1) Co-limitation of light and moisture; calculate `cb/cb_lightmax` and `cb/cb_moistmax`, and weight based on `DDMORT_CONST` (tunable global[?] parameter)
        - (2) "Leibig style"; limitation based on whether soil or light are more important (`cb/max(cb_lightmax, cb_moistmax)`)
    - No idea about this one. Need to do more reading.

### Lower priority

These are processes that I think will not have huge impacts, are farther-removed from the FoRTE project goals, or are things I simply don't understand very well (particularly all the micro-meteorology stuff).

- Branches (`IBRANCH_THERMO`)
    - Options:
        - (0) No branches in energy balance or biophysics
        - (1; default) Branches in energy balance, but not biophysics
        - (2) Branches in both
    - In theory, should be a minor component of the energy balance, so impact should be small. This is the hypothesis?
- Canopy roughness model (`ICANTURB`)
    - Options:
        - (0): Based on Leuning et al. (1995; TODO REF). Wind computed using "similarity theory" for top cohort, then extinguished with cumulative LAI.
        - (1, default) "Default ED2.1 scheme" (TODO: 0? Figure it out!), but using "zero-plane" displacement height.
        - (2) Massman (1997; TODO) using constant drag and no sheltering factor
        - (3) Also Massman, but with varying drag and sheltering within canopy
        - (4) Like 0, but calculate ground conductance following CLM (eq. 5.98-5.100)
    - No idea, except that (3) will have higher uncertainty than (2). Need to read more about this.
- Similarity theory model (for canopy roughness) (`ISFCLYRM`)
    - Computes `u*`, `T*`, etc.
    - Options:
        - (1) BRAMS default, based on Louis (1979 TODO). 
        - (2) Oncley and Dudhia (1995 TODO), based on MM5
        - (3) Beljaars and Holtslag (1991 TODO); Like (2), but alternative method for stable case that has more mixing
        - (4, default) CLM (2004; TODO: Check that this is same as 2013?). Like (2) and (3), but with special functions to deal with very stable cases.
    - Again, no idea. Need to read more about micro-meteorology and how it affects vegetation processes.
- Ground to canopy conductance (`IED_GRNDVAP`)
    - Options:
        - (0, default): Modified Lee Pielke (1992 TODO), with field capacity, but with beta factor without square (as in Noilhan and Planton 1989 TODO). Like original ED2 and LEAF-3.
        - (1-4) Mahfouf and Noilhan (1991 TODO), #1-4
        - (5) Combination of (1) (alpha) and (2) (soil resistance)
    - No idea. Again, need to do some micro-met reading.
- Tree-fall disturbance rate (`TREEFALL_DISTURBANCE_RATE`)
    - Options:
        - (0, default) No treefall disturbance
        - (> 0) Disturbance rate, in 1/years. Will create new patches.
        - (< 0) Also rate, but will be added to mortality (kill plants, but no new patches)
    - Reference value for this is 0.014 trees year^-1^ (note that because of ED's cohort approximation, number of individuals is continuous, not discrete)
    - See also `TIME2CANOPY`, minimum age for tree-fall disturbance.
- Growth respiration (`GROWTH_RESP_SCHEME`)
    - Options:
        - (0, default) Growth respiration is a tax on GPP, at PFT-specific rate. Treated as aboveground wood to canopy-airspace flux.
        - (1) Growth respiration calculated as in (0), but split into leaf, sapwood-a, sapwood-b, and root sources proportionally to biomass. No effect on overall C budget, but greater within-ecosystem flux resolution.
    - These should produce identical results, and if they don't, that means there is a bug!
    - If output is identical, we should set to (1), since we can always calculate total respiration otherwise.
- Storage respiration (`STORAGE_RESP_SCHEME`) -- same as growth
    - Note that this whole system is a _hack_, because there is no real "storage respiration" process. In PEcAn, we hard-code the relevant parameters to zero. An interesting question is whether we may be able to get better results with this hack enabled -- i.e. if it compensates a problem in some other process.

### Almost-certainly omit

These configurations are almost certainly irrelevant to the goals of this paper or will not apply at UMBS (e.g. grass or tropical species).
Mostly for reference.

- Grass (`IGRASS`) -- I'm not running with any grass PFTs.
- Phenology (`IPHEN_SCHEME`) -- Differences affect grasses and tropical species. Conifers are always evergreen, and hardwoods are always cold-deciduous (need to check that schemes are the same).
- Seed dispersal (`REPRO_SCHEME`) -- Only seems to matter for multi-site runs, with some toggles for tropical (drought-deciduous) species.
- Hydrodynamics (`PLANT_HYDRO_SCHEME`) -- Seems mostly tropics-focused? Look into the Xu et al. (2016; New Phyt.) paper.
- Fire disturbance (`INCLUDE_FIRE`) -- We know there weren't any fires at UMBS during this time, and more generally are interested in stability.
    - This could potentially be fun to play with in a later paper about disturbance regimes. Potentially, with Jackie Matthes?
- Anthropogenic disturbances (`IANTH_DISTURB`) -- We are currently looking at stable state, without disturbances. But, since this pulls values from a dataset, this could be one way to apply the FoRTE treatments.