### 2019-07-10 analysis/paper/submodel-rtm.Rmd
Add draft of RTM submodel descriptions

---
title: "Submodel descriptions: Radiative transfer"
author: Alexey Shiklomanov
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
stopifnot(
  requireNamespace("purrr", quietly = TRUE),
  requireNamespace("gsl", quietly = TRUE)
)
```

# Common features

## Scattering and backscattering of canopy elements

Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\nu$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:

$$ \nu = R + T $$
$$ \omega = \frac{R + T + \frac{1}{4} (R-T)(1 - \chi)^2}{2 (R+T)} $$

Both of these quantities are calculated independently for leaves and wood, and then averaged based on the relative area indices of leaves (LAI, $L$) and wood (WAI, $W$) within a canopy layer.

$$ w_{leaf} = \frac{L}{L + W}$$ 
$$ \nu = \nu_{leaf} w_{leaf} + \nu_{wood} (1 - w_{leaf}) $$
$$ \omega = \omega_{leaf} w_{leaf} + \omega_{wood} (1 - w_{leaf}) $$

## Directional scattering and leaf angle distribution

The directional extinction coefficient ($K(\psi)_i$) can be expressed as:

$$ K(\psi)_i = \frac{G(\psi)}{cos(\psi)} $$

where $G(\psi)$ describes the mean projection per unit leaf area (or "relative projected leaf area") in direction $\psi$.

The leaf angle distribution function used in ED2 is the same as the one used in CLM 4.5 [@clm45_note]:

$$ G(\psi) = \phi_1 + \phi_2 cos(\psi) $$
$$ \phi_1 = 0.5 - 0.633 \chi - 0.33 \chi^2 $$
$$ \phi_2 = 0.877 (1 - 2 \phi_1) $$

where $\chi$ is the _leaf orientation factor_ parameter, defined such that -1 is perfectly vertical leaves, 1 is perfectly horizontal leaves, and 0 is randomly distributed leaf angles.

```{r g-function, echo = FALSE, fig.cap = gfun_cap}
gfun_cap <- "Leaf projection as a function of leaf orientation factor."
phi1_f <- function(orient) {
  0.5 - 0.633 * orient - 0.33 * orient ^ 2
}
phi2_f <- function(orient, phi1 = phi1_f(orient)) {
  0.877 * (1 - 2 * phi1)
}
gfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  mu <- cos(theta_rad)
  phi1 + phi2 * mu
}
max_theta <- 89
orient <- c(-0.5, -0.25, 0, 0.25, 0.5)
colors <- colorRampPalette(c("red", "black", "blue"))(length(orient))
plot(c(-90, 90), c(0, 0.9), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(G(psi)))
legend("bottom", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
    orient, colors,
    ~curve(gfunction(x, .x), -max_theta, max_theta,
           col = .y, add = TRUE)
))
```

```{r kfunction, echo = FALSE, fig.cap = kcap}
kcap <- "K function"
kfunction <- function(theta, orient, theta_rad = theta * pi / 180) {
  gfunction(theta, orient, theta_rad) / cos(theta_rad)
}
plot(c(-70, 70), c(0.2, 1.8), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(K(psi)))
legend("top", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
  orient, colors,
  ~curve(kfunction(x, .x), -70, 70,
         col = .y, add = TRUE)
))
```

An additional quantity is the average inverse optical depth per unit leaf/wood area ($\bar\mu$):

$$
\bar\mu = \frac{1}{\phi_2}\left( 1 - \frac{\phi_1}{\phi_2} \ln\left( 1 + \frac{\phi_2}{\phi_1}\right) \right)
$$

```{r}
mu_bar <- function(orient) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  x <- 1 - phi1 / phi2 * log(1 + phi2 / phi1)
  out <- x / phi2
  out[orient == 0] <- 1
  out
}
curve(mu_bar(x), -0.5, 0.5,
      xlab = "Orient factor",
      ylab = expression(bar(mu)))
```

## Beam backscatter

The beam backscatter (or "upscatter") coefficient for direct radiation, $\beta_0$, is defined as:

$$
\beta_0 = a_s(\psi) \frac{1 + \bar\mu K(\psi)}{\bar\mu K(\psi)}
$$

where $a_s(\psi)$ is the single scattering albedo coefficient, defined as:

$$
a_s(\psi) = \frac{1}{2}
\frac{G(\psi)}{\cos\psi \phi2 + G(\psi)}
\left(
1 - 
\frac{\cos\psi \phi1}{\cos\psi + G(\psi)}
\ln\left(
\frac{\cos\psi \phi1 + \cos\psi \phi2 + G(\psi)}{\cos\psi \phi1}
\right)
\right)
$$

(Note that, for simplicity, $a_s$ here is equivalent to $\frac{a_s}{\omega}$ in @clm45_note equation 3.15, where $\omega$ is the leaf backscatter.)

## Exponential decay of direct radiation

In both the two-stream and multiple scatter schemes, without finite canopy radius, the transmissivity of a layer to direct radiation ($\tau(\psi)_i$) is defined as follows:

$$ \tau(\psi)_i = \exp(-K(\psi)_i L_i) $$

(Note that the inverse optical depth, $\mu_0$, in the two-stream calculation is just $K(\psi)^{-1}$).

## Matrix solution for diffuse fluxes

Both the two-stream and multiple-scatter canopy radiative transfer schemes are set up as the solutions to the following linear matrix equation:

$$ M \times X = Y $$

For $n$ canopy layers, $M$ is a $(2m+2, 2m+2)$ matrix and $X$ and $Y$ $(2m+2)$ vectors.
$X$ is the vector of fluxes (or values that can be converted to fluxes), formatted as follows:

$$
\begin{bmatrix}
F_{up,ground} \\
F_{down,1} \\
F_{up, 1} \\
... \\
F_{down,i} \\
F_{up,i} \\
... \\
F_{down,n} \\
F_{up,n} \\
F_{down,sky}
\end{bmatrix}
$$

$M$ and $Y$ consist of coefficients, which are defined differently in the two models.
Notably, in both cases $M$ is a sparse matrix, composed of zero elements everywhere except the diagonal and its immediate neighbors;
for example, for $n = 3$, $M$ looks like this:

$$
M = \begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 \\
m_{2,1} & m_{2,2} & m_{2,3} & 0 & 0 & 0 \\
0 & m_{3,2} & m_{3,3} & m_{4,3} & 0 & 0 \\
0 & 0 & m_{4,3} & m_{4,4} & m_{4,5} & 0 \\
0 & 0 & 0 & m_{5,4} & m_{5,5} & m_{5,6} \\
0 & 0 & 0 & 0 & 0 & 1 \\
\end{bmatrix}
$$

# Multiple scatter

All of this is based on @zhao_2005_multiple.

The $X$ and $Y$ vectors of the matrix problem for diffuse fluxes are defined as follows:

$$
X =
\begin{bmatrix}
SWuO_0 \\
SWd0_1 \\
SWu0_1 \\
... \\
SWd0_i \\
SWu0_i \\
... \\
SWd0_m \\
SWu0_m \\
SWd0_{m + 1}
\end{bmatrix}
$$

$$
Y = 
\begin{bmatrix}
S_0 a_{ground} \\
S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
... \\
S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
... \\
S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
SW_{sky}
\end{bmatrix}
$$

Here, $a_{ground}$ is the albedo of the ground under the canopy and $SW_{sky}$ is the incident shortwave hemispherical flux from the sky;
both are exogenous inputs to the model.
$S_i$ is the direct ("beam") radiation at layer $i$, and is calculated in a loop as follows:

$$ S_i = S_{i + 1} \tau(\psi)_i $$

with $S_{n+1}$ as the incident direct solar flux, an exogenous input.

$M$ is a sparse $(2n + 2)$ by $(2n + 2)$ matrix (where $m$ is the number of vertical canopy layers) that is zero everywhere except on the diagonal and its immediate neighbors.
For example, for $m = 2$:

For $i = 1,2,3,...n$, where $n$ is the number of canopy layers.

$$ m_{1,1} = m_{2n+2,2n+2} = 1 $$
$$ m_{2i,2i-1} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$
$$ m_{2i,2i} = -r_{i-1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i-1})(1 - \tau_{i-1}) $$
$$ m_{2i,2i+1} = 1 - r_{i-1} r_i (1 - \alpha_{i-1})(1 - \tau_{i-1})(1 - \alpha_i)(1 - \tau_i) $$
$$ m_{2i+1,2i} = 1 - r_i r_{i+1} (1 - \alpha_i)(1 - \tau_i)(1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i,2i} = -r_{i+1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i+1,2i+2} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$

The layer absorption ($\alpha$) and backscatter for direct ($r(\psi)_i$) and diffuse ($r_i$) light are defined as follows:

$$ \alpha_i = 1 - \nu $$
$$ r_i = \omega $$
$$ r(\psi)_i = \beta_0 $$

The _direct_ ("beam") transmittance fraction of canopy layer $i$ at solar zenith angle direction $\psi$ ($\tau(\psi)_i$) is defined as:

$$ \tau(\psi)_i = \exp\left( -K(\psi)_i L_i \right) $$

```{r int-direct, fig.cap = idcap}

idcap <- "Direct transmittance fraction, as a function of LAI"
tau_direct <- function(psi, L, orient) {
  exp(-kfunction(psi, orient) * L)
}
Lmax <- 5
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Direct tranmittance fraction")
legend("topright", c("-0.5", "0", "0.5"),
       col = c("red", "black", "blue"), lty = 1,
       title = "Orient factor")
legend("top", c("0", "30", "60"), lty = 1:3,
       title = "Zenith angle")
curve(tau_direct(0, x, 0), 0, Lmax, add = TRUE)
curve(tau_direct(0, x, -0.5), 0, Lmax, add = TRUE, col = "red")
curve(tau_direct(0, x, 0.5), 0, Lmax, add = TRUE, col = "blue")
curve(tau_direct(30, x, 0), 0, Lmax, add = TRUE, lty = 2)
curve(tau_direct(30, x, -0.5), 0, Lmax, add = TRUE, lty = 2, col = "red")
curve(tau_direct(30, x, 0.5), 0, Lmax, add = TRUE, lty = 2, col = "blue")
curve(tau_direct(60, x, 0), 0, Lmax, add = TRUE, lty = 3)
curve(tau_direct(60, x, -0.5), 0, Lmax, add = TRUE, lty = 3, col = "red")
curve(tau_direct(60, x, 0.5), 0, Lmax, add = TRUE, lty = 3, col = "blue")

```

where $L_i$ is the layer's leaf area index (technically, total area index), and $K(\psi)_i$ is the directional extinction coefficient (see "Directional scattering and leaf angle distribution").

For _diffuse_ radiation,
the interception fraction ($\tau_i$) is defined similarly except that it must be integrated across the hemisphere.
This is simplified by the assumption that incident diffuse radiation is isotropic.

$$
\tau_i = 2 \int_{0}^{\frac{\pi}{2}}
\exp \left( -K(\psi)_i L_i \right) \sin \psi \cos \psi d\psi
$$

For the $K(\psi)$ defined in "Directional scattering and leaf angle distribution", this integral has the following analytical solution:

$$
\tau_i = -\exp(-L_i (\phi_1 + \phi_2))
\left(
\phi_1^2 L_i^2 \exp(\phi_1 L_i) * E_i(-\phi_1 L_i) +
\phi_1 L_i - 1
\right)
$$

where $\phi_1$ and $\phi_2$ are defined as above, and $E_i$ is the exponential integral function.

```{r}

tau_diffuse <- function(L, orient) {
  phi1 <- phi1_f(orient)
  phi2 <- phi2_f(phi1 = phi1)
  ext_diff1 <- phi1 * L
  ext_diff2 <- phi2 * L
  -exp(-ext_diff1 - ext_diff2) *
    (ext_diff1^2 * exp(ext_diff1) * gsl::expint_Ei(-ext_diff1) +
       ext_diff1 - 1)
}
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
purrr::walk2(
  orient, colors,
  ~curve(tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y)
)

```

## Finite canopy radius

Crown area ($C_i$) is computed as a function of DBH.

$$ C_i = a DBH^b $$

where $a$ is 2.490154 and $b$ is 0.8068806 for all of the PFTs used in this study.

```{r}
dbh2ca <- function(dbh) {
  # default parameters for PFTs we use
  b1Ca <- 2.490154
  b2Ca <- 0.8068806
  b1Ca * dbh ^ b2Ca
}
curve(dbh2ca(x), 0, 50,
      xlab = "DBH (cm)",
      ylab = "Crown area index (unitless)")
```

The crown area index ($C_i$) affects the following terms in the following ways:

$$ \tau(\psi)_i = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$

$$
\tau_i = (1 - C_i) - 
C_i \exp(-\frac{L_i}{C_i} (\phi_1 + \phi_2))
\left(
\frac{\phi_1^2 L_i^2}{C_i^2} \exp(\phi_1 \frac{L_i}{C_i}) * E_i(-\phi_1 \frac{L_i}{C_i}) +
\phi_1 \frac{L_i}{C_i} - 1
\right)
$$

# Two-stream