### 2019-09-17 analysis/paper/submodel-rtm.Rmd
Add two-stream submodel code

---
title: "Description of radiative transfer submodels"
author: Alexey Shiklomanov
bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
stopifnot(
  requireNamespace("purrr", quietly = TRUE),
  requireNamespace("gsl", quietly = TRUE)
)
library(fortebaseline)
```

# Common features

## Scattering and backscattering of canopy elements

Following @clm45_note, forward- ($\nu$) and backscattering ($\omega$) of canopy elements (leaves or stems) are defined as a function of those elements' reflectance ($R$) and transmittance ($T$):

$$ \nu_x = R_x + T_x $$
$$ \omega_x = \frac{R_x + T_x + \frac{1}{4} (R_x-T_x)(1 - \chi)^2}{2 (R_x+T_x)} $$

Both of these quantities are calculated independently for leaves and wood, and then averaged based on the relative area indices of leaves (LAI, $L$) and wood (WAI, $W$) within a canopy layer.

$$ w_{leaf} = \frac{L}{L + W}$$ 
$$ \nu = \nu_{leaf} w_{leaf} + \nu_{wood} (1 - w_{leaf}) $$
$$ \omega = \omega_{leaf} w_{leaf} + \omega_{wood} (1 - w_{leaf}) $$

## Directional scattering and leaf angle distribution

The directional extinction coefficient ($K(\psi)_i$) can be expressed as:

$$ K(\psi)_i = \frac{G(\psi)}{cos(\psi)} $$

where $G(\psi)$ describes the mean projection per unit leaf area (or "relative projected leaf area") in direction $\psi$.

The leaf angle distribution function used in ED2 is the same as the one used in CLM 4.5 [@clm45_note]:

$$ G(\psi) = \phi_1 + \phi_2 cos(\psi) $$
$$ \phi_1 = 0.5 - 0.633 \chi - 0.33 \chi^2 $$
$$ \phi_2 = 0.877 (1 - 2 \phi_1) $$

where $\chi$ is the _leaf orientation factor_ parameter, defined such that -1 is perfectly vertical leaves, 1 is perfectly horizontal leaves, and 0 is randomly distributed leaf angles.

```{r g-function, echo = FALSE, fig.cap = gfun_cap}
gfun_cap <- "Leaf projection as a function of leaf orientation factor."
max_theta <- 89
orient <- c(-0.5, -0.25, 0, 0.25, 0.5)
colors <- colorRampPalette(c("red", "black", "blue"))(length(orient))
plot(c(-90, 90), c(0, 0.9), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(G(psi)))
legend("bottom", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
    orient, colors,
    ~curve(gfunction(x, .x), -max_theta, max_theta,
           col = .y, add = TRUE)
))
```

```{r kfunction, echo = FALSE, fig.cap = kcap}
kcap <- "K function"
plot(c(-70, 70), c(0.2, 1.8), type = "n",
     xlab = "Zenith angle (degrees)",
     ylab = expression(K(psi)))
legend("top", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
invisible(purrr::map2(
  orient, colors,
  ~curve(kfunction(x, .x), -70, 70,
         col = .y, add = TRUE)
))
```

An additional quantity is the average inverse optical depth per unit leaf/wood area ($\bar\mu$):

$$
\bar\mu = \frac{1}{\phi_2}\left( 1 - \frac{\phi_1}{\phi_2} \ln\left( 1 + \frac{\phi_2}{\phi_1}\right) \right)
$$

```{r}
curve(mu_bar(x), -0.5, 0.5,
      xlab = "Orient factor",
      ylab = expression(bar(mu)))
```

## Beam backscatter

The beam backscatter (or "upscatter") coefficient for direct radiation, $\beta_0$, is defined as:

$$
\beta_0 = a_s(\psi) \frac{1 + \bar\mu K(\psi)}{\bar\mu K(\psi)}
$$

where $a_s(\psi)$ is the single scattering albedo coefficient, defined as:

$$
a_s(\psi) = \frac{1}{2}
\frac{G(\psi)}{\cos\psi \phi_2 + G(\psi)}
\left(
1 - 
\frac{\cos\psi \phi_1}{\cos\psi + G(\psi)}
\ln\left(
\frac{\cos\psi \phi_1 + \cos\psi \phi_2 + G(\psi)}{\cos\psi \phi_1}
\right)
\right)
$$

(Note that, for simplicity, $a_s$ here is equivalent to $\frac{a_s}{\omega}$ in @clm45_note equation 3.15, where $\omega$ is the leaf backscatter.)

```{r single-scattering-albedo, fig.cap = ssa_cap}
ssa_cap <- paste("Single-scattering albedo as a function of ",
                 "radiation angle and leaf orientation factor.")
plot(c(-90, 90), c(0, 1.3), type = "n",
     xlab = expression("Zenith angle" ~ (psi * degree)),
     ylab = expression("Single-scattering albedo" ~ (a[s])))
curve(single_scatter_albedo(x, 0), -89, 89, add = TRUE)
curve(single_scatter_albedo(x, 0.25), -89, 89, col = "blue", add = TRUE)

curve(single_scatter_albedo(x, -0.25), -89, 89, col = "red", add = TRUE)
legend("topright", c("-0.25", "0", "0.25"), title = "Orient factor",
       col = c("red", "black", "blue"), lty = 1)
```

```{r beta0, fig.cap = beta0_cap}
beta0_cap <- paste(
  "Backscatter for direct radiation",
  "as a function of radiation angle",
  "and leaf orientation factor."
)
plot(c(-90, 90), c(0.4, 4), type = "n",
     xlab = expression("Zenith angle" ~ (psi * degree)),
     ylab = expression("Direct radiation backscatter" ~ (beta[0])))
curve(beta0(x, 0), -89, 89, add = TRUE)
curve(beta0(x, 0.25), -89, 89, col = "blue", add = TRUE)
curve(beta0(x, -0.25), -89, 89, col = "red", add = TRUE)
legend("topright", c("-0.25", "0", "0.25"), title = "Orient factor",
       col = c("red", "black", "blue"), lty = 1)
```

## Exponential decay of direct radiation

In both the two-stream and multiple scatter schemes, without finite canopy radius, the transmissivity of a layer to direct radiation ($\tau(\psi)_i$) is defined as follows:

$$ \tau(\psi)_i = \exp(-K(\psi)_i L_i) $$

(Note that the inverse optical depth, $\mu_0$, in the two-stream calculation is just $K(\psi)^{-1}$).

```{r int-direct, fig.cap = idcap}
idcap <- "Direct transmittance fraction, as a function of LAI"
Lmax <- 5
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Direct tranmittance fraction")
legend("topright", c("-0.5", "0", "0.5"),
       col = c("red", "black", "blue"), lty = 1,
       title = "Orient factor")
legend("top", c("0", "45"), lty = 1:2,
       title = "Zenith angle")
curve(tau_direct(x, 0, 0), 0, Lmax, add = TRUE)
curve(tau_direct(x, 0, -0.5), 0, Lmax, add = TRUE, col = "red")
curve(tau_direct(x, 0, 0.5), 0, Lmax, add = TRUE, col = "blue")
curve(tau_direct(x, 45, 0), 0, Lmax, add = TRUE, lty = 2)
curve(tau_direct(x, 45, -0.5), 0, Lmax, add = TRUE, lty = 2, col = "red")
curve(tau_direct(x, 45, 0.5), 0, Lmax, add = TRUE, lty = 3, col = "blue")
```

## Matrix solution for diffuse fluxes

Both the two-stream and multiple-scatter canopy radiative transfer schemes are set up as the solutions to the following linear matrix equation:

$$ M \times X = Y $$

For $n$ canopy layers, $M$ is a $(2m+2, 2m+2)$ matrix and $X$ and $Y$ $(2m+2)$ vectors.
$X$ is the vector of fluxes (or values that can be converted to fluxes), formatted as follows:

$$
\begin{bmatrix}
F_{up,ground} \\
F_{down,1} \\
F_{up, 1} \\
... \\
F_{down,i} \\
F_{up,i} \\
... \\
F_{down,n} \\
F_{up,n} \\
F_{down,sky}
\end{bmatrix}
$$

$M$ and $Y$ consist of coefficients, which are defined differently in the two models.
Notably, in both cases $M$ is a sparse matrix, composed of zero elements everywhere except the diagonal and its immediate neighbors;
for example, for $n = 3$, $M$ looks like this:

$$
M = \begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 \\
m_{2,1} & m_{2,2} & m_{2,3} & 0 & 0 & 0 \\
0 & m_{3,2} & m_{3,3} & m_{4,3} & 0 & 0 \\
0 & 0 & m_{4,3} & m_{4,4} & m_{4,5} & 0 \\
0 & 0 & 0 & m_{5,4} & m_{5,5} & m_{5,6} \\
0 & 0 & 0 & 0 & 0 & 1 \\
\end{bmatrix}
$$

# Multiple scatter

All of this is based on @zhao_2005_multiple.

The $X$ and $Y$ vectors of the matrix problem for diffuse fluxes are defined as follows:

$$
X =
\begin{bmatrix}
SWuO_0 \\
SWd0_1 \\
SWu0_1 \\
... \\
SWd0_i \\
SWu0_i \\
... \\
SWd0_m \\
SWu0_m \\
SWd0_{m + 1}
\end{bmatrix}
$$

$$
Y = 
\begin{bmatrix}
S_0 a_{ground} \\
S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
... \\
S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
... \\
S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
SW_{sky}
\end{bmatrix}
$$

Here, $a_{ground}$ is the albedo of the ground under the canopy and $SW_{sky}$ is the incident shortwave hemispherical flux from the sky;
both are exogenous inputs to the model.
$S_i$ is the direct ("beam") radiation at layer $i$, and is calculated in a loop as follows:

$$ S_i = S_{i + 1} \tau(\psi)_i $$

with $S_{n+1}$ as the incident direct solar flux, an exogenous input.

$M$ is a sparse $(2n + 2)$ by $(2n + 2)$ matrix (where $m$ is the number of vertical canopy layers) that is zero everywhere except on the diagonal and its immediate neighbors.
For example, for $m = 2$:

For $i = 1,2,3,...n$, where $n$ is the number of canopy layers.

$$ m_{1,1} = m_{2n+2,2n+2} = 1 $$
$$ m_{2i,2i-1} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$
$$ m_{2i,2i} = -r_{i-1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i-1})(1 - \tau_{i-1}) $$
$$ m_{2i,2i+1} = 1 - r_{i-1} r_i (1 - \alpha_{i-1})(1 - \tau_{i-1})(1 - \alpha_i)(1 - \tau_i) $$
$$ m_{2i+1,2i} = 1 - r_i r_{i+1} (1 - \alpha_i)(1 - \tau_i)(1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i,2i} = -r_{i+1} \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right] (1 - \alpha_{i+1})(1 - \tau_{i+1}) $$
$$ m_{2i+1,2i+2} = - \left[ \tau_i + (1 - \tau_i)(1 - \alpha_i)(1 - r_i) \right]$$

The _direct_ ("beam") transmittance fraction of canopy layer $i$ at solar zenith angle direction $\psi$ ($\tau(\psi)_i$) is defined above.

The layer absorption ($\alpha$) and backscatter for direct ($r(\psi)_i$) and diffuse ($r_i$) light are defined as follows:

$$ \alpha_i = 1 - \nu $$
$$ r_i = \omega $$
$$ r(\psi)_i = \beta_0 $$

For _diffuse_ radiation,
the interception fraction ($\tau_i$) is defined similarly except that it must be integrated across the hemisphere.
This is simplified by the assumption that incident diffuse radiation is isotropic.

$$
\tau_i = 2 \int_{0}^{\frac{\pi}{2}}
\exp \left( -K(\psi)_i L_i \right) \sin \psi \cos \psi d\psi
$$

For the $K(\psi)$ defined in "Directional scattering and leaf angle distribution", this integral has the following analytical solution:

$$
\tau_i = -\exp(-L_i (\phi_1 + \phi_2))
\left(
\phi_1^2 L_i^2 \exp(\phi_1 L_i) E_i(-\phi_1 L_i) +
\phi_1 L_i - 1
\right)
$$

where $\phi_1$ and $\phi_2$ are defined as above, and $E_i$ is the exponential integral function.

```{r}

plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
purrr::walk2(
  orient, colors,
  ~curve(ms_tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y)
)

```

# Two-stream

The $X$ and $Y$ vectors of the matrix problem for diffuse fluxes are defined as follows:

$$
X =
\begin{bmatrix}
SWuO_0 \\
SWd0_1 \\
SWu0_1 \\
... \\
SWd0_i \\
SWu0_i \\
... \\
SWd0_m \\
SWu0_m \\
SWd0_{m + 1}
\end{bmatrix}
$$

$$
Y = 
\begin{bmatrix}
S_0 a_{ground} () \\
S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
... \\
S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
... \\
S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
SW_{sky}
\end{bmatrix}
$$

Diffuse transmittance.

```{r}

plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
purrr::walk2(
  orient, colors,
  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
)

```

Here it is compared with the multi-scatter model.

```{r}

orient_sub <- c(-0.5, 0, 0.5)
colors_sub <- colors[c(1, 3, 5)]
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient_sub),
       title = "Orient factor",
       lty = "solid",
       col = colors_sub)
legend("top", c("multi-scatter", "two-stream"),
       title = "RTM",
       lty = c("dashed", "solid"))
purrr::walk2(
  orient_sub, colors_sub,
  ~curve(ms_tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
)
purrr::walk2(
  orient_sub, colors_sub,
  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
)

```

# Finite canopy radius

Crown area ($C_i$) is computed as a function of DBH.

$$ C_i = a DBH^b $$

where $a$ is 2.490154 and $b$ is 0.8068806 for all of the PFTs used in this study.

```{r}
dbh2ca <- function(dbh) {
  # default parameters for PFTs we use
  b1Ca <- 2.490154
  b2Ca <- 0.8068806
  b1Ca * dbh ^ b2Ca
}
curve(dbh2ca(x), 0, 50,
      xlab = "DBH (cm)",
      ylab = "Crown area index (unitless)")
```

In both models, the transmittance of direct radiation corrected for crown area index ($\tau(\psi)_i^*$) is calculated as follows:

$$ \tau(\psi)_i^* = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$

```{r}
tau_psi_star <- function(L, C, theta, orient) {
  k <- kfunction(theta, orient)
  1 - C + C * exp(-k * L / C)
}
curve(tau_psi_star(1, x, 0, 0), 0, 1,
      ylim = c(0, 1),
      xlab = "Crown area index", y = expression(tau(psi) ^ "*"))
curve(tau_psi_star(2, x, 0, 0), 0, 1, col = 2, add = TRUE)
curve(tau_psi_star(3, x, 0, 0), 0, 1, col = 3, add = TRUE)
curve(tau_psi_star(4, x, 0, 0), 0, 1, col = 4, add = TRUE)
```

The effect of the crown area index on diffuse transmittance varies by model.
In the **two-stream** model, the crown area index affects the average diffuse optical depth ($\bar{\mu}$) as follows:

$$
\bar{\mu}^* = \frac{-L_i}{\log(1 - C_i + C_i \exp( -L_i / \bar{\mu} ))}
$$

(Note that if $C_i = 1$, this simplifies to $\bar{\mu}^* = \bar{\mu}$).

In the **multiple-scatter** model, if we define the diffuse transmittance expression ($\tau_i$) above as a function of leaf area index, $T_i(L)$,
then the diffuse transmittance corrected for crown area ($\tau_i^*$) becomes:

$$
\tau_i^* = (1 - C_i) + C_i T_i\left( \frac{L_i}{C_i} \right)
$$

Comparing the diffuse radiation response of the two models:

```{r}

cai <- c(0.1, 0.25, 0.5, 1)
colors_sub <- c("red1", "red2", "red3", "black")
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("bottomleft", as.character(cai),
       title = "CAI",
       lty = "solid",
       col = colors_sub,
       inset = c(0, 0.2))
legend("bottomleft", c("MS", "2S"),
       title = "RTM",
       lty = c("dashed", "solid"))
purrr::walk2(
  cai, colors_sub,
  ~curve(ms_tau_diffuse(x, 0, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
)
purrr::walk2(
  cai, colors_sub,
  ~curve(ts_tau_diffuse(x, 0, lr = 0.056, lt = 0.045, cai = .x), 0, Lmax, add = TRUE, col = .y)
)

```

# Comparison

First, we generate some common inputs:

```{r, echo = TRUE}
ipft <- c(2, 1, 1)
L <- c(1, 1.5, 2)
p4 <- rrtm::prospect4(1.4, c(40, 60), 0.01, 0.01)
vis <- p4[(400:700) - 399, , ]
lr <- colMeans(vis[,,1])
lt <- colMeans(vis[,,2])

orient <- c(0.2, 0.25)

theta <- 15
S0_beam <- 0.8
S0_diff <- 1 - S0_beam
alb_ground <- 0.1
```

Then we run both models.

```{r}
ms_out <- multi_scatter(ipft, L, lr, lt, orient,
                        theta, S0_beam, S0_diff, alb_ground)
ts_out <- two_stream(ipft, L, lr, lt, orient,
                     theta, S0_beam, S0_diff, alb_ground)

plot(ms_out$light_level, type = "l")
lines(ts_out$light_level, col = 2)

plot(ms_out$light_diff_level, type = "l")
lines(ts_out$light_diff_level, col = 2)
```

# References